## ANOVA repeated measures

Perform a repeated measures ANOVA and get posthocs

**Libraries**
```{r libraries, message=FALSE}
if (!require('pacman')) install.packages('pacman'); library('pacman')
p_load(tidyverse, yarrr, readxl, stringr, afex) 
```



**Import data**  
We use the following [dataset](Data/08-Data_analysis/ANOVA_repeated_measures.xlsx).
```{r import_data}
#Import to variable "datos"
datos = read_excel("Data/08-Data_analysis/ANOVA_repeated_measures.xlsx") %>% 
  dplyr::mutate(ID = 1:n())

datos
```

***

**EXPERIMENT DESIGN**  
2 x 2 x 2 - Repeated measures  

* VI1 = ENDO / EXO
* VI2 = Temporal windows (bc - before change vs  ac - after change)
* VI3 = Percepts (h - horse vs m - morse)
 
***

**Prepare data**
```{r anova_prepare_data}
# Wide to Long
datos_long = datos %>% gather(Condicion, VD, 1:8)


# We extract conditions from Condicion column. Convert char variables to factor
datos_long = datos_long %>%
  mutate(VI1 = str_sub(datos_long$Condicion, 1,3)) %>%
  mutate(VI2 = 
           ifelse(VI1 == "END", 
                  str_sub(datos_long$Condicion, 5,6), 
                  str_sub(datos_long$Condicion, 4,5))) %>%
  mutate(VI3 = 
           ifelse(VI1 == "END", 
                  str_sub(datos_long$Condicion, 7,7), 
                  str_sub(datos_long$Condicion, 6,6))) %>%
  mutate(VI1 = as.factor(VI1)) %>%
  mutate(VI2 = as.factor(VI2)) %>%
  mutate(VI3 = as.factor(VI3)) %>%
  mutate(Condicion = as.factor(Condicion))

datos_long
```


Repeated measures ANOVA
```{r anova_afex_repeated_measures}
# AFEX PACKAGE
# https://cran.r-project.org/web/packages/afex/afex.pdf

a = aov_ez("ID", "VD", datos_long,
           within = c("VI1", "VI2", "VI3"))

# Show results
summary(a)

# Plot results
lsmip(a, VI1 ~ VI2 * VI3)
pirateplot(VD ~ VI1 * VI2 * VI3, datos_long)

```

Post-hocs
```{r anova_posthocs}
lsmeans(a, c("VI1", "VI2", "VI3"), contr = "pairwise")

lsmeans(a, "VI1", contr = "pairwise")
lsmeans(a, "VI2", contr = "pairwise")
lsmeans(a, "VI3", contr = "pairwise")

```