## Reliability

A continuación mostramos dos sistemas para calcular reliability. Alpha de Chronbach y Omega.

### Cronbach's Alpha

https://www.ncbi.nlm.nih.gov/pubmed/28557467?dopt=Abstract

** TODO **
* Items negatively correlated
* Alphadrop function para ordenar 


En R hay, al menos, dos funciones:

* `cronbach.alpha()` - `library(ltm)` - interfeeres with dplyr `select()`
* `alpha()` - `library(psych)`
  + We use `alpha()` -  Gives correlation information, etc.


* Procedure ** WIP: references for this procedure (?) **
1. **Delete** items with correlations (r.drop) < 0.25 one by one:
    + Eliminate the item with the smaller correlation `min(a$item.stats[["r.drop"]])`
    + Run reliability
    + Eliminate the next item with the smaller correlation (<0.25!) - **Leave at least 3 items!**
2. Reasonable Cronbach score: >0.6. Good score >0.7


* Cargamos librerias  
```{r alpha_libraries, message=FALSE}
if (!require('pacman')) install.packages('pacman'); library('pacman')
p_load(readr, psych, dplyr)
```

***
We use the following [dataset](Data/10-Pre_processing/Reliability.csv).

***  
```{r alpha_data_preparation}
temp = read_csv("Data/10-Pre_processing/Reliability.csv")
temp = temp[-1] 

#Convierte todo a integer
# temp = temp %>% dplyr::mutate_each(funs(as.integer))
temp = as.data.frame(sapply(temp, as.integer))

head(temp)

```


```{r alpha}
alpha(temp)
# temp2 = temp %>% select(-1, -OA.OA05.)
# alpha(temp2)
# temp2 = temp %>% select(-1, -OA.OA05., -OA.OA04.)
# alpha(temp2)
temp2 = temp %>% select(-1, -OA.OA05., -OA.OA04., -OA.OA02.)
alpha(temp2)

```


### Omega

* Cargamos librerias
```{r omega_libraries, message=FALSE}
if (!require('pacman')) install.packages('pacman'); library('pacman')
p_load(readr, psych, MBESS, dplyr, GPArotation)
# install.packages("MBESS", dependencies = TRUE)
```

***
We use the following [dataset](Data/10-Pre_processing/Reliability.csv).

***  
```{r omega_data_preparation}
temp = read_csv("Data/10-Pre_processing/Reliability.csv")
temp = temp[-1] 

#Convierte todo a integer
# temp = temp %>% dplyr::mutate_each(funs(as.integer))
temp = as.data.frame(sapply(temp, as.integer))

head(temp)

```

#### Method 1 - with Bootstrap ^1^

```{r omega_bootstrap, cache=TRUE}
# Use at least (?) B = 1000

# Execution time is non-trivial
set.seed(1)
ci.reliability(data = temp, type = "omega", conf.level = 0.95, interval.type = "bca", B = 1000)


```

#### Method 2 ^2^

```{r omega}

# See options
omega(temp)


```

### References

https://www.ncbi.nlm.nih.gov/pubmed/28557467?dopt=Abstract

^1^ http://onlinelibrary.wiley.com/doi/10.1111/bjop.12046/pdf

^2^ 

* https://www.researchgate.net/post/Cronbachs_alpha_vs_model_based_reliability_estimates_omega_and_others_is_it_worth_the_effort
    + http://personality-project.org/r/psych/HowTo/R_for_omega.pdf
    + Revelle, W. & Zinbarg, R.E. (2009) Coefficients alpha, beta, omega and the glb: comments on Sijtsma. Psychometrika. 74, 1, 145-154 (http://personality-project.org/revelle/publications/rz09.pdf)
    + Zinbarg, R.E., Revelle, W., Yovel, I., & Li. W. (2005). Cronbach's Alpha, Revelle's Beta, McDonald's Omega: Their relations with each and two alternative conceptualizations of reliability. Psychometrika. 70, 123-133. (http://personality-project.org/revelle/publications/zinbarg.revelle.pmet.05.pdf) 
    + Cronbach's alpha vs. model based reliability estimates (omega and others): is it worth the effort?: https://www.researchgate.net/post/Cronbachs_alpha_vs_model_based_reliability_estimates_omega_and_others_is_it_worth_the_effort
