---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Estadistica descriptiva{#Estadistica_descriptiva}

***
* Vamos a usar la siguiente [base de datos](Data/07_Descriptive_statistics/Descriptive_statistics.csv).

***

Cargamos librerias y leemos los datos
```{r descriptives-load, message=FALSE, warning=FALSE}
if (!require('findviews')) install.packages('findviews'); library('findviews')
if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
if (!require('readr')) install.packages('readr'); library('readr')

if (!require('gt')) install.packages('gt'); library('gt')
if (!require('gtsummary')) install.packages('gtsummary'); library('gtsummary')


datos = read_csv(here::here("Data/07_Descriptive_statistics/Descriptive_statistics.csv")); datos

```

## Gtsummaries

- Para mas informacion y ejemplos de codigo: https://themockup.blog/posts/2020-09-04-10-table-rules-in-r/

Create a simple descriptive table:

```{r gt1}

gtsummary::tbl_summary(datos,
                       by = Sexo,
                       missing = "ifany") %>% 
  gtsummary::add_n()

```


More complex:

Create a table for each Sex, combine the two and save to a file.

```{r gt2}

table1 = gtsummary::tbl_summary(datos %>% dplyr::filter(Sexo == 0) %>% dplyr::select(-Sexo),
                       by = Condition,
                       missing = "ifany",
                       type = list(Edad ~ 'categorical'),
                       statistic = list(Edad ~ "{n} ({p}%)")) %>% 
  gtsummary::add_n()


table2 = gtsummary::tbl_summary(datos %>% dplyr::filter(Sexo == 1) %>% dplyr::select(-Sexo),
                       by = Condition,
                       missing = "ifany",
                       type = list(Edad ~ 'categorical'),
                       statistic = list(Edad ~ "{n} ({p}%)")) %>% 
  gtsummary::add_n()


table_combined_Sexo = gtsummary::tbl_merge(list(table1, table2), tab_spanner = list("Females", "Males"))
table_combined_Sexo

# Save table
gtsave(as_gt(table_combined_Sexo), file = here::here("Resultados/table-combined-bysex.png"))


```


## Summary de datos
```{r descriptives-summary}
# datos %>% group_by(Condition)  %>% summarise(VD = mean(VD), sd = sd(VD))
    
datos %>% 
  group_by(Condition) %>% 
  summarise(sd = sd(VD), VD = mean(VD))

# datos %>% filter(Edad > 18) %>% group_by(Condition, Sexo)  %>% summarise(VD = mean(VD), stdev = sd(VD), VD_t = mean(VD_t), num = length(VD))
    
datos %>% 
  filter(Edad > 18) %>%
  group_by(Condition, Sexo) %>%
  summarise(num = length(VD), stdev = sd(VD), VD = mean(VD), VD_t = mean(VD_t))

# Todo en uno
datos %>%
  group_by(Condition) %>%
  summarise(num = length(VD), stdev = sd(VD), mean = mean(VD), max(VD), min(VD))

```


### Agrupamos por sujeto
```{r descriptives-group}
datos %>%
  group_by(ID) %>%
  summarise(mean = mean(VD))

```

## Descriptive tables latex

Using stargazer we can create tables showing summary statistics or a lm output. Here we use `type = "html"` to be able to show the table in this book. `type = "latex"` (default) shows the latex code.
```{r descriptives-tables, results='asis'}
if (!require('stargazer')) install.packages('stargazer'); library('stargazer')

stargazer(datos %>% as.data.frame(), type="html") 
```


## Visualize missing values

With Amelia
```{r descriptives-missing-amelia}
if (!require('Amelia')) install.packages('Amelia'); library('Amelia')

datos = datos %>% mutate(Sexo = ifelse(Edad == 37, NA, Sexo))

missmap(datos)

```

With ggplot and reshape2
```{r descriptives-missing-ggplot}
# With ggplot
# A function that plots missingness
if (!require('reshape2')) install.packages('reshape2'); library('reshape2')
if (!require('ggplot2')) install.packages('ggplot2'); library('ggplot2')

ggplot_missing <- function(x){
  
  x %>% 
    is.na %>%
    melt %>%
    ggplot(data = .,
           aes(x = Var2,
               y = Var1)) +
    geom_raster(aes(fill = value)) +
    scale_fill_grey(name = "",
                    labels = c("Present","Missing")) +
    theme_minimal() + 
    theme(axis.text.x  = element_text(angle=45, vjust=0.5)) + 
    labs(x = "Variables in Dataset",
         y = "Rows / observations")
}

ggplot_missing(datos)

```


```{r visualize-missing-naniar}

if (!require('naniar')) install.packages('naniar'); library('naniar')

# Add some missing in Edad
set.seed(10)
missing = rbinom(261, 1, 0.3)
datos$Edad = with(datos, ifelse(Edad >= 30 & missing == 1, NA, Edad))

# Visualize upset plot
  # Missing in Edad, Missing in Sexo, Missing in Sexo AND Edad
datos %>% 
  gg_miss_upset()



```



## Tutorial externo
* [Tutorial para colapsar datos](http://stanford.edu/~ejdemyr/r-tutorials/collapsing-data/)



## Findviews
Lanzar el siguiente comando para explorar visualmente los datos:

´findviews(datos)´

Ver pagina en [Github ](https://github.com/tsellam/findviews)
