## Association test

A.K.A. X square test of independence

**WIP - Alternatives**
```{r}
# Con datos de estudio Merlin
# prop.test(table(DB_all$Ahorro_Dico, DB_all$Futuro_d))
# chisq.test(table(DB_all$Ahorro_Dico, DB_all$Futuro_d))

# # More powerful
# Exact::exact.test(table(DB_all$Ahorro_Dico, DB_all$Futuro_d))
# BayesFactor::contingencyTableBF(table(DB_all$Ahorro_Dico, DB_all$Futuro_d), sampleType = "indepMulti", fixedMargin = "rows")

```

**Libraries**
```{r association_test_libraries, message=FALSE}
if (!require('pacman')) install.packages('pacman'); library('pacman')
p_load(BayesFactor, lsr)
```



**Import data - WIP**
We use the following [dataset](Data/08-Data_analysis/Association_test.csv).
```{r association_test_data_import}
#Import to variable "dataset"
dataset = read.csv("Data/08-Data_analysis/Association_test.csv")

#Show first 5 rows
head(dataset)
```

**Prepare data**
```{r association_test_data_preparation}
# Convert to factors
dataset$Accuracy_factor = as.factor(dataset$Accuracy)
dataset$Formato_factor = as.factor(dataset$Formato)
dataset$Problema_factor = as.factor(dataset$Problema)
```


### Frequentist [^1]
```{r association_test_frequentist}
# Accuracy_factor + Formato_factor
crosstab_Formato <- xtabs(~ Accuracy_factor + Formato_factor, dataset); crosstab_Formato
associationTest(formula = ~ Accuracy_factor + Formato_factor, data = dataset )

# Accuracy_factor + Problema_factor
crosstab_Problema <- xtabs(~ Accuracy_factor + Problema_factor, dataset); crosstab_Problema
associationTest(formula = ~ Accuracy_factor + Problema_factor, data = dataset )
```

### Bayesian [^2]
```{r association_test_bayesian}
# Accuracy_factor + Formato_factor
contingencyTableBF( crosstab_Formato, sampleType = "jointMulti" ) 

# Accuracy_factor + Problema_factor
contingencyTableBF( crosstab_Problema, sampleType = "jointMulti" ) 

```



### Results report - WIP
> TODO: Use an example where we have both a null a and a significant finding

#### Frequentist - WIP
Pearson’s X^2^ revealed a significant association between species and choice (X^2^ = 10.7, p = .01): robots appeared to be more likely to say that they prefer flowers, but the humans were more likely to say they prefer data.

#### Bayesian
We ran a Bayesian test of association (see Gunel & Dickey, 1974) using version 0.9.12-2 of the BayesFactor package (Morey & Rouder, 2015) using default priors and a joint multinomial sampling plan. The resulting Bayes factor of 3.65 to 1 in favour of the null hypothesis indicates that there is moderately strong evidence for the non-independence of Problem and accuracy.


### References
[^1]: http://health.adelaide.edu.au/psychology/ccs/docs/lsr/lsr-0.5.pdf#page=368  
[^2]: http://health.adelaide.edu.au/psychology/ccs/docs/lsr/lsr-0.5.pdf#page=571  
