# Pre-processing

## RT filtering

### Function to filter by RT

Here you can see an example on how to use a function to filter by RT (log filtering version below).


***
We use the following [dataset](Data/10-Pre_processing/RT_filtering.csv).

***

```{r}
library(tidyverse)

# RT Filtering function
Filter_FUN <- function(DB, Column, Low_PCT = 0.05, High_PCT = 0.95) {
  
  args = as.list(match.call())
  
  # Creamos Column_LOG
  DB = DB %>% mutate(Column_LOG = log(eval(args$Column, DB)))
  lowcut = quantile(DB$Column_LOG, Low_PCT, na.rm = T)
  highcut = quantile(DB$Column_LOG, High_PCT, na.rm = T)
  
  # Returns DB_Filtered
  DB_Filtered <<- subset(DB, DB$Column_LOG > lowcut & DB$Column_LOG < highcut) 
  
  # return(DB_Filtered)
  cat("**** Datos filtrados a partir de la columna", args$Column, ": ", nrow(DB) - nrow(DB_Filtered), "/", nrow(DB), "filas eliminadas ***")
  
}
```

After creating the function, we read the DB and apply the filtering. A dataframe called DB_Filtered is created with the filtered data. 

```{r}

dataset = read_csv("Data/10-Pre_processing/RT_filtering.csv")
dataset

Filter_FUN(dataset, Time)
DB_Filtered
```


### n SD from the median (MAD) [^1]
> Probably should NOT use n SD from mean

***
We use the following [dataset](Data/10-Pre_processing/RT_filtering.csv).

***

```{r}

dataset = read.csv("Data/10-Pre_processing/RT_filtering.csv")
head(dataset)

```

* Code
```{r}
#SD criteria
SD_filter = 2.5

dataset_filtered = dataset %>% 
  filter(Time < (median(Time) + SD_filter * sd(Time)) & Time > (median(Time) - SD_filter * sd(Time)))

# dataset_filtered = dataset[dataset$Time < (median(dataset$Time) + SD_filter * sd(dataset$Time) )  & dataset$Time > (median(dataset$Time) - SD_filter * sd(dataset$Time) ), ]
```

* Plots
```{r rt_filtering_MAD_plots}
summary(dataset$Time)
summary(dataset_filtered$Time)

par(mfrow = c(1,2))
    #
    plot(density(dataset$Time), main = "Valid reaction time density")
    plot(density(dataset_filtered$Time), main = "Valid reaction time density")
    
par(mfrow = c(1,1))
```



### Gamma law regression analyses [^2]

***
We use the following [dataset](Data/10-Pre_processing/RT_filtering.csv).

***

```{r}
#Import to variable "dataset"
dataset = read.csv("Data/10-Pre_processing/RT_filtering.csv")
head(dataset)
```

* Code
```{r}
#Gamma
glm1 = (glm(Time ~ Formato, data = dataset, family = Gamma))
summary(glm1)

#Gausian
glm2 = (glm(Time ~ Formato, data = dataset))
summary(glm2)

```



### Logarithms

Log filtering may not be the best option available [^2]

***
We use the following [dataset](Data/Pre-processing/RT_Filtering_data_file.csv).

***
```{r}
#Import to variable "dataset"
dataset = read.csv("Data/10-Pre_processing/RT_filtering.csv")
head(dataset)
```

* Code
```{r}

#SET low and high PERCENTILES
    lowcutquant = 0.05
    highcutquant = 0.95
    
    # Add log-rt
    # dataset_filtered = data_long_RAW
    dataset$logrt <- log(dataset$Time)
    
    # Take correct answers
        # WE USE CORRECT ANSWERS TO SET TIME LIMITS!!!
        # BETTER USE dataset?
    corr = subset(dataset, dataset$Accuracy == 1)
    lowcut = quantile(corr$logrt, lowcutquant)
    highcut = quantile(corr$logrt, highcutquant)
    dataset_filtered <- subset(dataset, dataset$logrt > lowcut & dataset$logrt < highcut)
```

* Plots
```{r rt_filtering_LOG_plots}
plot(density(dataset_filtered$logrt), main = "Log reaction time density")
    
par(mfrow = c(1,2))

    #
    plot(density(dataset$Time[dataset$Time < 10]), main = "Density of small reaction times")
    plot(density(dataset_filtered$Time[dataset_filtered$Time < 10]), main = "Density of small reaction times")

    #
    plot(density(dataset$Time), main = "Valid reaction time density")
    plot(density(dataset_filtered$Time), main = "Valid reaction time density")
    
par(mfrow = c(1,1))

    #
    summary(dataset$Time)
    summary(dataset_filtered$Time)

```

### References  
[^1]: Leys, C., Ley, C., Klein, O., Bernard, P. & Licata, L. Detecting outliers: Do not use standard deviation around the mean, use absolute deviation around the median. J. Exp. Soc. Psychol. 49, 764–766 (2013).
[^2]: [See **Serban C. Musca** post](https://www.researchgate.net/post/For_decision_times_do_you_first_take_the_logarithm_and_then_exclude_extreme_outliers_or_the_other_way_around_And_how_do_you_define_outliers)