# Experimental design

## Tamaño de muestra y efecto detectado

En esta simulación se muestra como con tamaños de muestra pequeños, se sobreestima el tamaño de efecto real. Como se puede ver en el codigo, solo se muestran los estudios simulados donde la correlacion es significativa. Ver el [blog de Jim Grange](https://jimgrange.wordpress.com/2017/03/06/low-power-effect-sizes/) para una explicación más detallada.
```{r sample_size_calculation}
# https://jimgrange.wordpress.com/2017/03/06/low-power-effect-sizes/
#------------------------------------------------------------------------------
rm(list = ls())

set.seed(50)

# function for generating random draws from multivariate distribution
# n = number of draws; p = number of variables
# u = mean of each variable; s = SD of each variable
# corMat = correlation matrix
mvrnorm <- function(n, p, u, s, corMat) {
  Z <- matrix(rnorm(n * p), p, n)
  t(u + s * t(chol(corMat)) %*% Z)
}

#------------------------------------------------------------------------------

#------------------------------------------------------------------------------
### simulation setup

# declare simulation parameters
means <- c(100, 600)
sds <- c(20, 80)
n_sims = 1000

# declare correlation coefficient & generate cor. matrix
cor <- 0.3
cor_mat <- matrix(c(1, cor, 
                    cor, 1), nrow = 2, ncol = 2, byrow = TRUE)

# sample sizes to simulate
sample_sizes <- c(10, 20, 30, 50, 85, 170, 500, 1000)

# number of simulations

# create variable to store data in
final_data <- matrix(nrow = n_sims, ncol = length(sample_sizes))
colnames(final_data) <- sample_sizes
#------------------------------------------------------------------------------

#------------------------------------------------------------------------------
### simulation execution

for(i in 1:length(sample_sizes)){
  for(j in 1:n_sims){
    
    # get the experiment data
    sim_data <- mvrnorm(sample_sizes[i], p = 2, u = means, s = sds, 
                        corMat = cor_mat)
    
    # perform the correlation
    sim_cor <- cor.test(sim_data[, 1], sim_data[, 2], method = "pearson")
    
    # if the correlation is significant, store the effect size
    if(sim_cor$p.value < 0.05){
      final_data[j, i] <- as.numeric(round(abs(sim_cor$estimate), 3))
    }
  }
}
#------------------------------------------------------------------------------

#------------------------------------------------------------------------------
## draw the plot
boxplot(final_data, na.action=na.omit, xlab = "Simulated Sample Size", 
        ylab = "Measured Effect Size")
abline(h = cor, col = "red", lty = 2, lwd = 3)

legend("topright", "True Effect Size", cex = 1.5,  lty = 2, bty="n", 
       col = "red", lwd = 3)
#------------------------------------------------------------------------------
```



## Power analysis

Tenemos un conjunto de funciones que nos permiten hacer power analysis para distintos tipos de diseños:  

* pwr.t.test  
* pwr.t2n.test  
* pwr.anova.test  

En cada uno de ellos podemos modificar varios parámetros. El parámetro que dejemos sin especificar será el que se calculará:  

* k = groups  
* n = participants (in each group)  
* f = effect size  
* sig.level = sig level  
* power = power  



**Libraries**
```{r power_analysis_libraries, message=FALSE, warning=FALSE}
if (!require('pacman')) install.packages('pacman'); library('pacman')
p_load(pwr)
```


**Code**
```{r power_analysis_code}

#T-test
pwr.t.test(n = , d = 0.8, sig.level = 0.05, power = 0.95, type = c("two.sample"))

#T-test 
pwr.t2n.test(n1 = 150, n2 = 75, d = , sig.level = 0.05, power = 0.8)


#Anova 
pwr.anova.test(k = 3, n = , f = .25, sig.level = 0.05, power = .8)


#Effect size that we can detect
pwr.anova.test(k = 6, n = 42, sig.level = 0.05, power = .95)


#Sample size needed to detect an effect of f=0.5
pwr.anova.test(k = 2, f = 0.5, sig.level = 0.05, power = .95)

```

### Longitudinal designs

Calculate sample size for longitudinal designs

**Libraries**
```{r power_analysis_longitudinal_libs, message=FALSE, warning=FALSE}
if (!require('pacman')) install.packages('pacman'); library('pacman')
p_load(powerMediation)
```


**Code**
```{r power_analysislongitudinal}
ssLong(es = 0.5, #	effect size of the difference of mean change.
       rho = 0.5, # correlation coefficient between baseline and follow-up values within a treatment group.
       alpha = 0.05, 
       power = 0.95)
```

